---
title: "GAPMINDER | LIFE EXPECTANCIES 2007"
author: "Adrián Calvo-Valderrama"
format: dashboard
theme: lux
---

```{r}
if(!require('pacman')) install.packages('pacman')

pacman::p_load(tidyverse,
               plotly,
               gapminder,
               reactable,
               bslib,
               bsicons,
               shiny,
               rnaturalearth,
               countrycode,
               htmltools,
               reactable)
```

```{r, include=FALSE}

gap_2007 <-
  gapminder %>% 
  filter(year == 2007)

highest_lifexp_country <- 
gap_2007 %>% 
  arrange(-lifeExp) %>% 
  head(1) %>% 
  pull(country)

highest_lifexp_value <- 
gap_2007 %>% 
  arrange(-lifeExp) %>% 
  head(1) %>% 
  pull(lifeExp) %>% 
  round(1)


lowest_lifexp_country <- 
gap_2007 %>% 
  arrange(lifeExp) %>% 
  head(1) %>% 
  pull(country)

lowest_lifexp_value <- 
gap_2007 %>% 
  arrange(lifeExp) %>% 
  head(1) %>% 
  pull(lifeExp) %>% 
  round(1)


average_lifexp_value <- 
gap_2007 %>% 
  summarise(
    weighted_mean = sum(lifeExp * pop)/ sum(pop)) %>% 
  round(1)


country_shapes <- rnaturalearth::ne_countries()

ggplot(country_shapes) +
  geom_sf()


country_shapes

gap_2007 <- gap_2007 %>% 
  mutate(country_code = countrycode(country, 
                                    "country.name", 
                                    "iso3c"))



setdiff(gap_2007$country_code, country_shapes$adm0_a3)


gap_map <- 
  left_join(country_shapes, gap_2007, by = c("adm0_a3" = "country_code")) %>% 
  mutate(tooltip_label = paste(country, round(lifeExp, 1),
                               sep = ": ")) %>% 
  ggplot() + 
  geom_sf(aes(fill = lifeExp, text = tooltip_label)) +
  theme_void() + 
  theme(legend.position = "none")

gap_map_ggplotly <- 
  ggplotly(gap_map, tooltip = "text")


top_20_countries <-
  gap_2007 %>% 
  arrange(-lifeExp) %>% 
  head(20) %>% 
   mutate(tooltip_label = paste(country, round(lifeExp, 1),
                               sep = ": ")) %>% 
  ggplot(aes(y = reorder(country, lifeExp),
             x = lifeExp,
             fill = lifeExp,
             text = tooltip_label)) +
  geom_col() + 
  geom_text(aes(label = round(lifeExp, 1)),
            nudge_x = -10,
            colour = "white") +
  labs(y = "Country",
       x = "Life Exp") + 
  theme(legend.position = "none")

top_20_countries_ggplotly <-
  ggplotly(top_20_countries, tooltip = "text")

```

# HOME 


## Row 1 {height=25%}

```{r}
value_box(
  title = "Highest Life Expectancy",
  value = paste(highest_lifexp_value, "years"),
  showcase = bsicons::bs_icon("heart"),
  theme = value_box_theme(bg = "#518fd6"),
  p(paste0("(", highest_lifexp_country, ")"))
)
```

```{r}
value_box(
  title = "Lowest Life Expectancy",
  value = paste(lowest_lifexp_value, "years"),
  showcase = bsicons::bs_icon("thermometer-low"),
  theme = value_box_theme(bg = "#214773"),
  p(paste0("(", lowest_lifexp_country, ")"))
)
```

```{r}
value_box(
  title = "Average Life Expectancy",
  value = paste(average_lifexp_value, "years"),
  showcase = bsicons::bs_icon("intersect"),
  theme = value_box_theme(bg = "#3f71ab")
)
```



## Row 2 {height=75%}

### {width=70%}

```{r title = "Map of Countries by Life Expectancy"}
gap_map_ggplotly
```


### {width=30%}


```{r title = "Top 20 Countries by Life Expectancy"}
top_20_countries_ggplotly
```


# DOWNLOAD DATA

The data used in this dashboard is shown below and can be downloaded as a CSV.


```{r}



library(htmltools)

htmltools::browsable(
  tagList(reactable(gapminder,
      elementId = "gapminder-table",
      searchable = TRUE,
      filterable = TRUE),
      

  tags$button("Download as CSV", 
              onclick = "Reactable.downloadDataCSV('gapminder-table')"),
  )
)

```

# ABOUT

This data comes from the r `gapminder` package and is originally sourced from the Gapminder Foundation.

Gapminder Foundation is a non-profit venture registered in Stockholm, Sweden, that promotes sustainable global development and achievement of the United Nations Millennium Development Goals by increased use and understanding of statistics and other information about social, economic, and environmental development at local, national, and global levels.

Gapminder was founded in 2005 by Ola Rosling, Anna Rosling Rönnlund, and Hans Rosling. The name Gapminder was derived from the "Mind the Gap" warning messages on the London Underground.

An example of one of Hans Rosling's videos is shown belowÑ

<iframe width="560" height="315" src="https://www.youtube.com/embed/hVimVzgtD6w?si=v44yz4iYO0GsGxX1" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>

